### 2023-09-20 paper/paper.tex
Changes from papeeria

\documentclass[10pt,twocolumn,letterpaper]{article}
%\usepackage[review]{cvpr}      % To produce the REVIEW version
%\usepackage{cvpr}              % To produce the CAMERA-READY version
\usepackage[pagenumbers]{cvpr} % To force page numbers, e.g. for an arXiv version
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage[pagebackref,breaklinks,colorlinks]{hyperref}
\usepackage{hyperref} % Import the hyperref package
\usepackage[capitalize]{cleveref}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}
\Crefname{table}{Table}{Tables}
\crefname{table}{Tab.}{Tabs.}
\def\cvprPaperID{00000} % *** Enter the CVPR Paper ID here
\def\confName{CVPR}
\def\confYear{2023}
\begin{document}
	\title{Enhancing LiDaR semantic segmentation using model soups }
	\author{ Omar Wasfy\\
	{\tt\small omar.wasfy@alexu.edu.eg}
\and
Marwan Torki\\
{\tt\small mtorki@alexu.edu.eg}
\and
Sohier Basiony\\
{\tt\small sohier.basiony@alexu.edu.eg}
}
\maketitle
\begin{abstract}
	In this paper, we present a pioneering approach for model soups \cite{wortsman2022model,dansereau2023model}, named "Iterative Scan Greedy Model Soup," and apply it to the domain of LiDAR semantic segmentation. using this technique, we augment the state-of-the-art open source code for 2DPASS \cite{yan20222dpass} semantic segmentation and demonstrate its efficacy on both the semanticKitti dataset and the nuscenes dataset. Remarkably, our approach achieves improvements in Mean Intersection Over Union (MIoU) without any increase in prediction time, all while retaining the original model structure.	Our contributions in this work are described in two points: Firstly, we successfully extend the application of model soups to LiDAR semantic segmentation, showcasing their adaptability and potential impact on the domain. Second, we introduce an optimized and efficient version of the existing greedy soup technique, further enhancing the overall performance of the approach. Through comprehensive experiments and rigorous evaluations, our results highlight the significance of the iterative scan greedy model soup in advancing LiDAR semantic segmentation, opening up new possibilities for enhanced perception systems in autonomous driving and related fields.
\end{abstract}
\section{Introduction}
\label{sec:intro}
In recent years, semantic segmentation of LiDAR point clouds has emerged as a critical area of research, especially to advance the scene understanding capabilities of self-driving cars.\cite{hu2019randla,yan2021sparse, SparseConv} The methods for semantic segmentation of LiDAR point clouds can be categorized based on their input modalities into two main groups:
Single Modality LiDAR-Based Input: These methods depend only on LiDAR data as input. However, due to the sparse nature of LiDAR input, such single-modality-only approaches often face challenges in accurately segmenting complex scenes.\cite{chen2017deeplab,chen2017rethinking,song2017semantic,huang2019ccnet,yan2021sparse,xu2020squeezesegv3,zhu2021cylindrical,tang2020searching,zheng2022beyond,zheng2021box}
Multiple-Modality LiDAR-Based and Camera-Based Input: These methods use both LiDAR-based and camera-based inputs. By fusing information from both modalities, these techniques attempt to exploit the strengths of each sensor to achieve more robust semantic segmentation.\cite{zhuang2021perception,el2019rgb,vora2020pointpainting}
The open-source model "2D Priors Assisted Semantic Segmentation" (2DPASS)\cite{yan20222dpass} utilizes multiple modalities during the training process to enhance the prediction capabilities of a single-modality model. However, after training, only the single modality (LiDAR) is retained, making it a more practical solution in the real world deployment t.
In the context of model ensembling, Model soups present an efficient approach for improving model performance without modifying the underlying model structure. This technique involves creating an ensemble of models with uniform structures and achieving performance gains akin to traditional ensembling methods without incurring additional prediction costs.
Building on the foundation of 2DPASS \cite{yan20222dpass}, we propose a novel enhancement to the approach. We introduce a procedure where multiple checkpoints are saved during the training process, and subsequently, a new round of optimization is performed using model soups techniques. By iteratively applying model soups, we aim to further enhance the predictive capabilities of the LiDAR-based single-modality model.
In this paper, we present our iterative scan greedy model soup approach, which leverages the strengths of both 2DPASS \cite{yan20222dpass} and model soups. We demonstrate the effectiveness of our method by conducting extensive experiments on LiDAR semantic segmentation tasks. The results showcase the potential of our approach to achieve notable performance improvements without altering the original model architecture and without incurring additional prediction time.
The contributions of this work are as follows. \\ 
  \begin{enumerate}
\item Our work involves the integration of model soups with 2DPASS \cite{yan20222dpass}, offering a fresh perspective to enhance LiDAR-based single-modality semantic segmentation models. To the best of our knowledge, no previous research has explored the application of the model souping technique to the task of LiDAR modality segmentation.
\item The validation of our iterative scan greedy model soup approach through various experiments demonstrates its substantial value in improving semantic segmentation, particularly for enhancing scene understanding in the context of self-driving cars.
\end{enumerate}
\section{Related Work}
\subsection{ Single-Sensor Methods
}
\subsubsection{Camera-Based Methods:}
These methods aim to predict pixel-wise labels for 2D images, with FCN \cite{long2015fully} being a pioneering architecture in this field. Recent advancements have explored multi-scale feature learning, dilated convolution, and attention mechanisms to achieve significant improvements. However, camera-only methods suffer from limitations in depth sensing and low-light conditions.\cite{he2016deep,chen2017rethinking,chen2017deeplab,lin2016efficient,zhao2017pyramid,wang2018understanding,huang2019ccnet,yuan2018ocnet}
\subsubsection{LiDAR-Based Methods:}
These methods process LiDAR point clouds directly. Point-based methods, like PointNet \cite{qi2016pointnet}, utilize per-point Multi-Layer Perceptron (MLP)\cite{rosenblatt1958perceptron} While these methods extract local features from point clouds, they can be inefficient due to time-consuming sampling and grouping algorithms.\cite{thomas2019kpconv,qi2017pointnet++,wang2019dynamic, PointConv,liu2019relation,Thomas_2019_ICCV,hua2018pointwise,yan2020pointasnl,zhao2021point,engel2021point,yan2020pointasnl,zhao2021point,engel2021point}
\subsubsection{Projection-based methods}
These methods project LiDAR point clouds onto 2D pixels, allowing the use of traditional CNNs. However, such methods suffer from information loss, leading to lower segmentation accuracy \cite{lawin2017deep, boulch2017unstructured, tatarchenko2018tangent,wu2018squeezeseg,wu2019squeezesegv2,liong2020amvnet}.
\subsubsection{Voxel-based frameworks}
These methods using SparseConv \cite{SparseConv}, have become popular due to their efficiency. These methods store only nonempty voxels in a Hash table and perform convolution operations more efficiently. Recent studies have utilized SparseConv \cite{SparseConv} to design powerful network architectures, but still face challenges to achieve greater segmentation accuracy \cite{zhou2020cylinder3d,cheng20212}.
\subsection{Multi-Sensor Methods}
These methods aim to combine information from both cameras and LiDAR, taking advantage of the strengths of both modalities. 
Approaches such as RGBAL \cite{el2019rgb} convert RGB images to polar-grid mapping representations and use early and mid-level fusion strategies. 
PointPainting \cite{vora2020pointpainting}  improves LiDAR network performance by projecting segmentation logits from images into the LiDAR space. 
PMF \cite{zhuang2021perception} explores the collaborative fusion of two modalities in camera coordinates. However, these methods require multi-sensor inputs during both training and inference, making them computationally intensive and less practical for real-world applications.\cite{krispel2020fuseseg,el2019rgb,meyer2019sensor,vora2020pointpainting}
\subsection{Cross-modal Knowledge Transfer}
Knowledge distillation \cite{hinton2015distilling}, initially was for model compression, has been extended to transfer knowledge between different models through the alignment of feature representations.\cite{ba2013deep,chen2017learning,zagoruyko2016paying,srinivas2018knowledge,gupta2016cross,wang2019efficient,yuan2018rgb,liu20213d,zhao2020knowledge}
Recently, knowledge distillation has been applied to transfer priors across different modalities, using extra 2D images during training to improve inference performance. 
Various approaches have been explored, such as 2D-assisted pre-training  \cite{liu2021learning}, 
inflating 2D convolution kernels to 3D \cite{xu2021image2point}, 
and well-designed teacher-student frameworks \cite{yuan2022x}.
In this work, we introduce a novel model soup \cite{wortsman2022model,dansereau2023model} technique applicable to all the above methods because of its inherent model-agnostic feature. 
This technique improves the results without changing the model architecture or incurring extra processing time. By leveraging the complementary information from 2D images, our method improves the segmentation accuracy of LiDAR point clouds in a cross-modal setup. We believe that this approach can be a valuable addition to the existing methods for multisensor semantic segmentation and also in other machine learning-based methods.
\section{ Approach }
Model souping \cite{wortsman2022model,dansereau2023model} is an ensemble learning technique that averages the weights of multiple models to create a single, more accurate, and robust model. This approach can help mitigate overfitting and improve generalization performance. Model souping is different from traditional ensemble learning techniques in that it averages the weights of models, rather than their predictions. This can help to improve the diversity of the ensemble and reduce the risk of overfitting. Model souping can be computationally expensive and requires more training data than traditional ensemble learning techniques. However, it can offer significant performance improvements over traditional methods.
\begin{table}[!htb]
\centering\
\small
\begin{tabular}{p{3.5cm} p{4cm}}
		\toprule
		\textbf{Method} & \textbf{Description} \\
		\midrule
		Best ValAcc Model & Choose model with highest validation accuracy ValAcc($\theta_i$) \\
		Ensemble & Combine predictions of models parameterized by $\theta_i$ \\
		Uniform Soup & Average parameter values $\theta_i$ \ \cite{wortsman2022model} \\
		Greedy Soup & Greedy approach for the model ensemble \cite{wortsman2022model}\\
		Learned Soup & Learn optimized ensemble of models \cite{wortsman2022model} \\
		Pruned Soups & Select subset of models for ensemble \cite{dansereau2023model} \\
		Iterative Uniform Greedy Soup Alg. 1 & Iterative Greedy Soup with Uniform Weights \\
        Greedier Iterative Uniform Greedy Soup Alg. 2 & same as Alg. 1 except to restart picking from the first checkpoint in the list after each ingredient is added \\
        Greedier Soup Alg. 3& same as Greedy Soup except for restarting picking from the first checkpoint in the list after each gradient is added. No duplicate picking allowed\\
        \\
        Greedier Pruned Soup Alg.4 & same as pruned soups with the exception of restarting removing checkpoints from the first checkpoint in the list after each ingredient is removed \\
		\bottomrule
	\end{tabular}
	\caption{Summary of Methods}
\end{table}
\begin{algorithm}[!htb]
\caption{Iterative Uniform Greedy Soup}
\label{alg:algorithm1}
\textbf{Input}: weights of Potential soup ingredients ${\theta_1, ..., \theta_k}$ (optionally sorted in descending order of \textbf{ValAcc}($\theta_i$))\\
\textbf{Parameter}:epochs\\
\begin{algorithmic}[1] %[1] enables line numbers
\STATE soupList = emptyList
\STATE bestValAcc = 0
\FOR{epoch=1 to epochs}
\STATE addedCheckpoints $\leftarrow$ 0
\FOR{i=1 to k}
\STATE potentialSoupList $\leftarrow$ AddToSoupList($\theta_i$)
\IF {\textbf{ValAcc}(potentialSoupList) $>=$ bestValAcc}
\STATE bestValAcc = \textbf{ValAcc}(potentialSoupList)
\STATE soupList $\leftarrow$ potentialSoupList
\STATE addedCheckpoints $\leftarrow$ addedCheckpoints + 1
\ENDIF
\ENDFOR
\IF{addedCheckpoints $>=$ 1}
\STATE BREAK
\ENDIF
\COMMENT{Early stopping}
\ENDFOR
\STATE \textbf{return} weights of the final soup $\theta_{soupList}$
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[!htb]
\caption{Greedier Iterative Uniform Greedy Soup}
\label{alg:algorithm2}
\textbf{Input}: weights of Potential soup ingredients ${\theta_1, ..., \theta_k}$ (optionally sorted in descending order of \textbf{ValAcc}($\theta_i$))\\
\textbf{Parameter}:epochs\\
\begin{algorithmic}[1] %[1] enables line numbers
\STATE soupList = emptyList
\STATE bestValAcc = 0
\WHILE{TRUE}
\FOR{epoch=1 to epochs}
\STATE addedCheckpoints $\leftarrow$ 0
\FOR{i=1 to k}
\STATE potentialSoupList $\leftarrow$ AddToSoupList($\theta_i$)
\IF {\textbf{ValAcc}(potentialSoupList) $>=$ bestValAcc}
\STATE bestValAcc = \textbf{ValAcc}(potentialSoupList)
\STATE soupList $\leftarrow$ potentialSoupList
\STATE addedCheckpoints $\leftarrow$ addedCheckpoints + 1
\STATE BREAK
\ENDIF
\ENDFOR
\ENDFOR
\COMMENT{Early stopping}
\IF{addedCheckpoints $>=$ 1}
\STATE BREAK
\ENDIF
\COMMENT{Early stopping}
\ENDWHILE
\STATE \textbf{return} weights of the final soup $\theta_{soupList}$
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[!htb]
\caption{Greedier Iterative Uniform Greedy Soup}
\label{alg:algorithm3}
\textbf{Input}: weights of Potential soup ingredients ${\theta_1, ..., \theta_k}$ (optionally sorted in descending order of \textbf{ValAcc}($\theta_i$))\\
\textbf{Parameter}:epochs\\
\begin{algorithmic}[1] %[1] enables line numbers
\STATE soupList = emptyList
\STATE bestValAcc = 0
\WHILE{TRUE}
\STATE addedCheckpoints $\leftarrow$ 0
\FOR{i=1 to k}
\IF{i IN soupList}
\STATE CONTINUE
\ENDIF
\STATE potentialSoupList $\leftarrow$ AddToSoupList($\theta_i$)
\IF {\textbf{ValAcc}(potentialSoupList) $>=$ bestValAcc}
\STATE bestValAcc = \textbf{ValAcc}(potentialSoupList)
\STATE soupList $\leftarrow$ potentialSoupList
\STATE addedCheckpoints $\leftarrow$ addedCheckpoints + 1
\STATE BREAK
\ENDIF
\ENDFOR
\IF{addedCheckpoints $>=$ 1}
\STATE BREAK
\ENDIF
\COMMENT{Early stopping}
\ENDWHILE
\STATE \textbf{return} weights of the final soup $\theta_{soupList}$
\end{algorithmic}
\end{algorithm}
\begin{algorithm}[!htb]
\caption{Greedier Pruned Soup}
\label{alg:algorithm4}
\textbf{Input}: weights of Potential soup ingredients ${\theta_1, ..., \theta_k}$ (optionally sorted in ascending order of \textbf{ValAcc}($\theta_i$))\\
\textbf{Parameter}:epochs\\
\begin{algorithmic}[1] %[1] enables line numbers
\STATE removedSoupList = emptyList
\STATE bestValAcc = 0
\WHILE{TRUE}
\STATE removedCheckpoints $\leftarrow$ 0
\FOR{i=1 to k}
\IF{i IN soupList}
\STATE CONTINUE
\ENDIF
\STATE potentialSoupList $\leftarrow$ RemoveFromSoupList($\theta_i$)
\IF {\textbf{ValAcc}(potentialSoupList) $>=$ bestValAcc}
\STATE bestValAcc = \textbf{ValAcc}(potentialSoupList)
\STATE removedSoupList $\leftarrow$ potentialSoupList
\STATE removedCheckpoints $\leftarrow$ removedCheckpoints + 1
\STATE BREAK
\ENDIF
\ENDFOR
\IF{addedCheckpoints $>=$ 1}
\STATE BREAK
\ENDIF
\COMMENT{Early stopping}
\ENDWHILE
\STATE \textbf{return} weights of the final soup $\theta_{soupList}$
\end{algorithmic}
\end{algorithm}
\section{Experiments}
\subsection{Experiment setup}
we have collected 100 checkpoints during training 2DPASS \cite{yan20222dpass} on both  Semantickitti \cite{behley2019semantickitti}. and Nuscenes \cite{caesar2020nuscenes}  datasets.
We have applied our modified greedy soup technique
also we have conducted a comparison between collected checkpoints  class-wise and average miou performance using tta with number of votes = 12 and showed the improvement of using tta against non-tta in semantic Kitti validation set for collected checkpoints while training   
\subsection{Results}.
\begin{table}[!htb]
\centering
\small
\footnotesize
	\caption{Semantic Kitti Performance Metrics for Selected Different Objects/Models}
	\begin{tabular}{{p{2cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm} p{0.5cm}}}
		\toprule
		Objects/Models & 1 ~\ref{row:1} & 2 ~\ref{row:2} & 9 ~\ref{row:9} & 10 ~\ref{row:10} & 7 ~\ref{row:7} & 5 ~\ref{row:5} & 4 ~\ref{row:4} \\
		\midrule
		car & 96.12 & 96.98 & 96.83 & 97.26 & 96.44 & 97.26 & 96.51 \\
		bicycle & 49.04 & 53.02 & 52.55 & 55.25 & 49.73 & 54.42 & 49.75 \\
		motorcycle & 73.68 & 76.69 & 76.31 & 78.72 & 71.89 & 76.28 & 72.36 \\
		truck & 79.11 & 92.5 & 90.75 & 90.79 & 84.94 & 95.34 & 85.22 \\
		bus & 64.24 & 70.85 & 71.38 & 72.08 & 68.06 & 74.69 & 69.22 \\
		person & 73.66 & 77.98 & 78.35 & 80.74 & 75.08 & 78.83 & 75.09 \\
		bicyclist & 87.8 & 91.41 & 92.35 & 93.55 & 88.77 & 92.97 & 89.16 \\
		motorcyclist & 0.19 & 0 & 0.06 & 0.14 & 0 & 0 & 0 \\
		road & 92.48 & 93.96 & 93.24 & 94.31 & 92.43 & 93.9 & 92.34 \\
		parking & 45.01 & 50.93 & 50.75 & 53.77 & 44.03 & 47.45 & 43.36 \\
		sidewalk & 78.84 & 81.53 & 80.08 & 82.03 & 78.95 & 81.56 & 78.81 \\
		other-ground & 4.81 & 7.32 & 8.44 & 12.09 & 5.2 & 7.19 & 5.44 \\
		building & 90.8 & 91.9 & 92.21 & 93 & 91.42 & 92.54 & 91.5 \\
		fence & 61.4 & 66.35 & 68.27 & 71.61 & 63.61 & 69.1 & 63.92 \\
		vegetation & 88.28 & 89.4 & 88.37 & 89.11 & 88.38 & 89.59 & 88.54 \\
		trunk & 70.71 & 73.17 & 71.19 & 72.23 & 70.98 & 72.62 & 70.93 \\
		terrain & 75.11 & 76.69 & 74.64 & 75.72 & 75.41 & 77.38 & 75.92 \\
		pole & 58.8 & 63.94 & 63.92 & 66.06 & 59.73 & 64.25 & 59.9 \\
		traffic-sign & 50.23 & 53.16 & 53.46 & 54.26 & 50.33 & 51.99 & 50.13 \\
		\midrule
		val/mIoU & 65.28 & 68.83 & 68.59 & 70.14 & 66.07 & 69.34 & 66.22 \\
		val/acc & 88.87 & 89.88 & 89.36 & 90.01 & 89.03 & 90.04 & 89.09 \\
		\bottomrule
	\end{tabular}
\end{table}
\begin{table}[!htb]
	\centering
	\footnotesize
	\caption{Nuscenes Performance Metrics for Different Objects/Models}
	\begin{tabular}{{p{3cm} p{1cm} p{1cm} p{1cm} p{1cm}}}
		\toprule
		Objects/Models & 1 ~\ref{row:1} & 9 ~\ref{row:9} & 10 ~\ref{row:10} & 4 ~\ref{row:4} \\
		\midrule
		movable\_object.barrier & 74.59 & 75.4 & 77.86 & 74.67 \\
		vehicle.bicycle & 36.27 & 42.48 & 50.37 & 36.28 \\
		vehicle.bus.bendy & 93.85 & 95.11 & 95.91 & 94.73 \\
		vehicle.car & 92.07 & 90.71 & 92.12 & 91.45 \\
		vehicle.construction & 53.95 & 57.39 & 61.38 & 55.08 \\
		vehicle.motorcycle & 80.88 & 85.65 & 87.97 & 81.72 \\
		human.pedestrian.adult & 78.15 & 78.65 & 81.84 & 78.21 \\
		movable\_object.trafficcone & 55.68 & 61.48 & 66.91 & 56.34 \\
		vehicle.trailer & 70.63 & 65.86 & 73.21 & 71.14 \\
		vehicle.truck & 86.77 & 86.04 & 89.27 & 87.55 \\
		flat.driveable\_surface & 96.04 & 96.08 & 96.68 & 96.11 \\
		flat.other & 71.75 & 73.02 & 74.74 & 73.04 \\
		flat.sidewalk & 73.4 & 73.15 & 75.12 & 73.76 \\
		flat.terrain & 73.82 & 73.84 & 75.56 & 74.28 \\
		static.manmade & 87.71 & 87.44 & 88.6 & 87.74 \\
		static.vegetation & 85.78 & 85.58 & 86.84 & 86.11 \\
		\midrule
		val/mIoU & 75.71 & 76.74 & 79.65 & 76.14 \\
		val/acc & 61.80 & 61.76 & 62.22 & 61.87 \\
		\bottomrule
	\end{tabular}
\end{table}
\begin{table}[!htb]
\centering
\caption{Validation results on SemanticKitti MIoU}
\label{tab:validation_results}
\begin{tabular}{|l|c|}
\hline
\textbf{Method} & \textbf{MIoU} \\
\hline
1 best individual ~\ref{row:1} & 65.28 \\
\hline
2 best individual (tta=12) ~\ref{row:2} &  68.83 \\
\hline
3 Algo1 no early stopping ~\ref{row:3} & 66.22 \\
\hline
4 Algo1 ~\ref{row:4}& 66.22 \\
\hline
5 Algo1 (tta=12) ~\ref{row:5} & 69.34 \\
\hline
6 Algo1 (tta=24) ~\ref{row:6} & 69.34 \\
\hline
7 Greedy Soup ~\ref{row:7}  & 66.07 \\
\hline
8 Greedy Soup (tta=12) ~\ref{row:8} &  69.21 \\
\hline
9 2DPASS github checkpoint ~\ref{row:9} & 68.59 \\
\hline
10 2DPASS  github checkpoint (tta=12) ~\ref{row:10} & 70.14 \\
\hline
11 Algo4~\ref{row:11} & 66.14 \\
\hline
12 Pruned Soup ~\ref{row:11} & [?] \\
\hline
13 Algo2 ~\ref{row:12} & 65.67 \\
\hline
14 Algo 3~\ref{row:13} & 66.05 \\
\hline
\end{tabular}
\end{table}
\begin{table}[!htb]
\centering
\caption{Testing results on SemanticKitti MIoU}
\label{tab:testing_results}
\begin{tabular}{|l|c|}
\hline
\textbf{Method} & \textbf{MIoU} \\
\hline
1 best individual ~\ref{row:1} &  62.82 \\
\hline
2 best individual (tta=12) ~\ref{row:2} &  65.87 \\
\hline
3 Algo1 no early stopping  ~\ref{row:3} & 62.99 \\
\hline
4 Algo1  ~\ref{row:4}& 63.18 \\
\hline
5 Algo1 (tta=12) ~\ref{row:5} & 65.70 \\
\hline
5 Algo1 (tta=24)  ~\ref{row:6} & 65.61 \\
\hline
7 Greedy Soup ~\ref{row:7}  & 63.17 \\
\hline
8 Greedy Soup (tta=12) ~\ref{row:8} &  65.66 \\
\hline
9 2DPASS github checkpoint ~\ref{row:9} & 65.30 \\
\hline
10 2DPASS  github checkpoint (tta=12) ~\ref{row:10} & 67.87 \\
\hline
\end{tabular}
\end{table}
\begin{table}[!htb]
\footnotesize
\caption{Explanation of methods in Table~\ref{tab:testing_results}}
\label{tab:explanation}
\begin{tabular}{|l|l|}
\hline
\textbf{Abbreviation} & \textbf{Explanation} \\
\hline
1 & Best individual checkpoint   \label{row:1}\\
\hline
2 & Best individual checkpoint with TTA =12 \label{row:2}\\
\hline
3 & Iterative Uniform Greedy Soup without early stopping \label{row:3}\\
\hline
4 & Iterative Uniform Greedy Soup   \label{row:4} \\
\hline
5 & Iterative Uniform Greedy Soup  TTA=24 \label{row:5} \\
\hline
6 & Iterative Uniform Greedy Soup   TTA=12 \label{row:6} \\
\hline
7 &  Greedy soup \label{row:7} \\
\hline
8 &  Greedy soup TTA=12 \label{row:8} \\
\hline
9 & 2DPASS Github best-provided checkpoint \label{row:9}  \\
\hline
10 & 2DPASS Github best-provided checkpoint TTA=12 \label{row:10} \\
\hline
11 & Greedier Pruned Soup \label{row:11} \\
\hline
12 & Pruned Soup \label{row:12} \\
\hline
13 & Greedier Iterative Uniform Greedy Soup \label{row:13} \\
\hline
14 & Greedier Greedy Soup \label{row:14} \\
\hline
\end{tabular}
\end{table}
\begin{enumerate}
\item Using Test-Time Augmentation (TTA), Greedy Soup shows slightly lower performance compared to the best individual checkpoint of the training. Specifically, with Greedy Soup using TTA=24, the performance is 0.04 lower, and with TTA = 12, it is 0.1 lower.
\item We can observe that Greedy Soup performs better without Test-Time Augmentation, which is a form of ensembling. This result contrasts with the validation set, indicating a possible overfitting pattern. During the attempt to assemble the best combination of ingredients to improve mIoU validation, there might have been a tendency to slightly overfit the validation data set. Interestingly, Greedy Soup's model performs better than the best checkpoint in testing with TTA=1, suggesting an improvement in generalization capacity and indicating that Greedy Soup's model has been enhanced without overfitting.
\end{enumerate}
\subsection{Ablation study}
\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\textwidth]{photos/semantickitti_soup.png}
    \caption{Enhaced Greedy Soup improves 1\% on best checkpoint }
\end{figure}
\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\textwidth]{photos/nuscenes_soup.png}
    \caption{Enhanced Greedy Soup improves 0.5\% on best checkpoint }
\end{figure}
\section{Conclusion}
We introduce a new method named "Iterative Scan Greedy Model Soup" that we have applied to LiDAR semantic segmentation. By harnessing this innovative approach, we improve an advanced open source 2DPASS \cite{yan20222dpass} semantic segmentation code and show its effectiveness on the semantic Kitti and Nuscenes datasets. Significantly, our technique achieves remarkable enhancements in Mean Intersection Over Union (MIoU) metrics, all while keeping the prediction time unchanged.
\bibliographystyle{plain} % Choose a bibliography style
\bibliography{egbib} % Specify the .bib file name without the extension
\end{document}